Blum axioms and nondeterministic computation of functions

### 2016-04-20 paper.tex
paper.tex: Simplify criticism of Goldreich's definition.

\begin{definition}[Goldreich's definition]
    Let $\bot$ be some special symbol not in $\{0, 1\}^*$.
    (This symbol will represents ``don't know''.)
    A nondeterministic machine $M$ computes the function $f$ if,
    when processing the input $x$,
    both the following conditions hold \cite[p.~168]{Goldreich2008}:
    \begin{itemize}
        \item Every branch of $M$ halts
            and outputs either $f(x)$ or $\bot$.
        \item At least one branch of $M$ halts with $f(x)$ on the tape.
    \end{itemize}
\end{definition}

The extra symbol $\bot$ sidesteps the problems of a branch looping forever.
Now,
every branch is required to halt for the value to be defined.
Thus,
we can convert a nondeterministic Turing machine to a deterministic machine
that compute the same function,
albeit with an exponential slowdown
--- the deterministic machine just need to simulate all branches until completion,
to actually be sure every branch halts;
if some branch do not halt,
then the simulating machine will not halt either,
but the function is not defined in this case,
so this behavior is correct.

Also, a deterministic machine functions exactly the same way under Goldreich's definition,
so we can convert mechanically
between the numbering corresponding to this definition
and a standard numbering for deterministic machines.
Therefore,
to show the universal machine theorem and the $S_{mn}$ theorem,
we can first convert the machine in question to a deterministic machine
and use their theorems.
Thus, this definition comply with the requirements to being an acceptable Gödel numbering,

To establish a complexity measure,
we can count the number of steps of a branch with most steps,
and leave it undefined if some branch loops forever.
--- this is the same method
we use to evaluate the time complexity of a nondeterministic decider.
Leaving undefined if a branch loops forever guarantees the first Blum axiom is satisfied.
And, to decide the predicate of the second axiom,
we can simply run all the branches of the machine for the specified number of steps.
Every branch has to halt within that number of steps,
and some branch must take exactly that number,
for the predicate to be true.

So, Goldreich's definition defines an acceptable Gödel numbering
and we can form a complexity measure that satisfies Blum axioms.
But, again,
we have trouble with the ``exponential speedup'' requirement.

Take, for instance,
the characteristic function of the satisfiability problem.
We can create a machine that guesses an assignment
and write $1$ if it is satisfying and write $\bot$ if it isn't.
This computes in linear time the function $f$, defined by
\begin{equation*}
    f(\varphi) = \begin{cases}
        1, & \text{if $\varphi$ is satisfiable.} \\
        \text{undefined}, & \text{if $\varphi$ is unsatisfiable.} \\
    \end{cases}
\end{equation*}
However,
this function falls short of being exactly $\SAT$'s characteristic function:
$f(\varphi)$ ought to be zero for unsatisfiable formulas $\varphi$.
Although we cannot prove a strong assertion as
``if this works then $\P = \NP$'',
we can come close.

>>>>

\begin{definition}[Goldreich's definition]
    Let $\bot$ be some special symbol not in $\{0, 1\}^*$.
    (This symbol will represents ``don't know''.)
    A nondeterministic machine $M$ computes the function $f$ if,
    when processing the input $x$,
    both the following conditions hold \cite[p.~168]{Goldreich2008}:
    \begin{itemize}
        \item Every branch of $M$ halts
            and outputs either $f(x)$ or $\bot$.
        \item At least one branch of $M$ halts with $f(x)$ on the tape.
    \end{itemize}
\end{definition}
The extra symbol $\bot$ sidesteps the problems of a branch looping forever.
This allows us to mechanically convert nondeterministic machines
under Goldreich's definition
to deterministic machines via simulation
--- the deterministic machine just need to simulate all branches until completion,
to actually be sure every branch halts;
if some branch do not halt,
then the simulating machine will not halt either,
but the function is not defined in this case,
so this behavior is correct.

Thus, we only enumerate computable functions.
to show the universal machine theorem and the $S_{mn}$,
we can simply first convert the machine in question to a deterministic machine
and use their theorems;
thus, this definition yields an acceptable Gödel numbering.
And, by counting the number of steps of the deepest branch,
we have a Blum complexity measure.

So, Goldreich's definition defines an acceptable Gödel numbering
and we can form a complexity measure that satisfies Blum axioms.
But, again,
we have trouble with the ``exponential speedup'' requirement.
For instance,
a machine trying to solve the satisfiability problem
would correctly return $1$ for a satisfiable instance,
but no branch can write a $0$ alone
because it cannot be sure that instance is unsatisfiable
--- so, the function would be undefined for unsatisfiable formulas.

That is, unless $\NP = \coNP$.
